Star-galaxy Classification Using Deep Convolutional Neural Networks

### 2016-10-05 paper/main.tex
remove colors and strikeout

Using the CasJobs server, we randomly select a total of 65,000 sources,
which are either stars or galaxies.
\changed{
In this work, we exclude objects that clearly are neither stars nor galaxies.
Most of the excluded objects are QSOs or quasars.
Quasars appear as point sources, rather than resolved sources similar to
galaxies,
and many of them have one or more saturated pixels in the images.
However, unlike any known stars, their spectra show strong and broad emission
lines.
Quasars are also different from galaxies because of their intrinsic variability
on a wide range of time scales, which may be due to variation in the accretion
rate or instabilities of the accretion disk around the black hole
\citep{popovic2012photocentric}.
Thus, many studies exclude quasars in the binary star-galaxy classification
scheme~\citep[\eg][]{vasconcellos2011decision,Fadely2012}.
Expanding the historical star-galaxy classification problem to include
additional classes, \eg \textit{nsng} (neither star nor galaxy),
may have advantages~\citep{ball2006robust}, and we plan to present
the results of this multi-class problem in a future paper.
}

@@

In the remaining layers, we still follow VGGNet and add many
$3\times3$ convolutions (with zero-padding of size 1 pixel).
Note that with the padding of 1 pixel for $3\times3$ convolutional layers,
the spatial resolution will be preserved after convolution.
Such preservation of spatial resolution allows us to build relatively deep
networks, as shown in Table~\ref{table:hyperparamters}.
The main contribution of VGGNet is in showing that the depth plays an important
role in good performance.
In our case, we start with four convolutional layers and progressively add more
layers, while monitoring the validation loss;
we stop at eight 
\changed{
convolutional
}
layers after we find that adding more layers
leads to worse performance.
We conjecture that a greater depth and hence larger number of parameters
lead to overfitting in our case.

@@

To compare the performance of ConvNets with
machine learning algorithms that use standard photometric features,
we use a machine learning framework called
Trees for Probabilistic Classifications (TPC).
TPC is a parallel, supervised machine learning algorithm
that uses prediction trees and random forest 
techniques~\citep{breiman1984classification, breiman2001random}
to produce a star-galaxy classification.
A complete description of TPC is beyond the scope of this paper, and
we refer the reader to
\cite{carrascokind2013tpz} and \cite{kim2015hybrid} for more details.
While other random forest implementations exist, we have chosen TPC, because
it is tailored specifically for astronomical use~\citep{carrascokind2013tpz};
it has been tested for astronomical use cases, including
photometric redshift estimation~\citep{sanchez2014photometric}
and star-galaxy classification~\citep{kim2015hybrid};
\sout{
it adds an extra layer of randomness by perturbing the measured attributes by
their uncertainties;
}
and it uses parallelism to handle large data sets on distributed memory systems.

@@

\begin{figure*}
  \centering
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/galaxy_original.pdf}
    \caption{Input (5 bands$\times44\times44$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/galaxy_conv11.pdf}
    \caption{Layer 1 (32 maps$\times40\times40$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/galaxy_conv21.pdf}
    \caption{Layer 3 (64 maps$\times20\times20$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/galaxy_conv31.pdf}
    \caption{Layer 6 (128 maps$\times10\times10$)}
  \end{subfigure}
  \caption{
    (a) A sample $44\times44$ RGB image of a galaxy in the CFHTLenS data set.
    The RGB image is created by mapping R $\rightarrow i$ band magnitude,
    G $\rightarrow r$ band magnitude, and B $\rightarrow g$ band magnitude.
    (b) Activations on the first convolutional layer when a $5\times44\times44$
    image is fed into the network.
    (c) Activations on the third convolutional layer.
    (d) Activations on the sixth convolutional layer.
    Each image in 
\changed{
(b), (c), and (d)
}
is a feature map corresponding to the
    output for one of the learned features.
    }
  \label{fig:galaxy_activations}
\end{figure*}

>>>>

Using the CasJobs server, we randomly select a total of 65,000 sources,
which are either stars or galaxies.
In this work, we exclude objects that clearly are neither stars nor galaxies.
Most of the excluded objects are QSOs or quasars.
Quasars appear as point sources, rather than resolved sources similar to
galaxies,
and many of them have one or more saturated pixels in the images.
However, unlike any known stars, their spectra show strong and broad emission
lines.
Quasars are also different from galaxies because of their intrinsic variability
on a wide range of time scales, which may be due to variation in the accretion
rate or instabilities of the accretion disk around the black hole
\citep{popovic2012photocentric}.
Thus, many studies exclude quasars in the binary star-galaxy classification
scheme~\citep[\eg][]{vasconcellos2011decision,Fadely2012}.
Expanding the historical star-galaxy classification problem to include
additional classes, \eg \textit{nsng} (neither star nor galaxy),
may have advantages~\citep{ball2006robust}, and we plan to present
the results of this multi-class problem in a future paper.

@@

In the remaining layers, we still follow VGGNet and add many
$3\times3$ convolutions (with zero-padding of size 1 pixel).
Note that with the padding of 1 pixel for $3\times3$ convolutional layers,
the spatial resolution will be preserved after convolution.
Such preservation of spatial resolution allows us to build relatively deep
networks, as shown in Table~\ref{table:hyperparamters}.
The main contribution of VGGNet is in showing that the depth plays an important
role in good performance.
In our case, we start with four convolutional layers and progressively add more
layers, while monitoring the validation loss;
we stop at eight convolutional layers after we find that adding more layers
leads to worse performance.
We conjecture that a greater depth and hence larger number of parameters
lead to overfitting in our case.

@@

To compare the performance of ConvNets with
machine learning algorithms that use standard photometric features,
we use a machine learning framework called
Trees for Probabilistic Classifications (TPC).
TPC is a parallel, supervised machine learning algorithm
that uses prediction trees and random forest 
techniques~\citep{breiman1984classification, breiman2001random}
to produce a star-galaxy classification.
A complete description of TPC is beyond the scope of this paper, and
we refer the reader to
\cite{carrascokind2013tpz} and \cite{kim2015hybrid} for more details.
While other random forest implementations exist, we have chosen TPC, because
it is tailored specifically for astronomical use~\citep{carrascokind2013tpz};
it has been tested for astronomical use cases, including
photometric redshift estimation~\citep{sanchez2014photometric}
and star-galaxy classification~\citep{kim2015hybrid};
and it uses parallelism to handle large data sets on distributed memory systems.

@@

\begin{figure*}
  \centering
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/galaxy_original.pdf}
    \caption{Input (5 bands$\times44\times44$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/galaxy_conv11.pdf}
    \caption{Layer 1 (32 maps$\times40\times40$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/galaxy_conv21.pdf}
    \caption{Layer 3 (64 maps$\times20\times20$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/galaxy_conv31.pdf}
    \caption{Layer 6 (128 maps$\times10\times10$)}
  \end{subfigure}
  \caption{
    (a) A sample $44\times44$ RGB image of a galaxy in the CFHTLenS data set.
    The RGB image is created by mapping R $\rightarrow i$ band magnitude,
    G $\rightarrow r$ band magnitude, and B $\rightarrow g$ band magnitude.
    (b) Activations on the first convolutional layer when a $5\times44\times44$
    image is fed into the network.
    (c) Activations on the third convolutional layer.
    (d) Activations on the sixth convolutional layer.
    Each image in (b), (c), and (d) is a feature map corresponding to the
    output for one of the learned features.
    }
  \label{fig:galaxy_activations}
\end{figure*}
